MIRSVM: Multi-Instance Support Vector Machine with Bag Representatives

### 2017-06-12 ArticleRevision/article/article.tex
Edits
> Re-formatting text. Removed "procedure(s)", and edited the text so the bottom paragraph is on the same page.
> Fixed error with figure labeling
> Next: begin reformatting to els

Algorithm~\ref{alg:mirsvm} shows the procedures for training the multi-instance representative SVM classifier and obtaining the optimal representatives from each bag.  During training, the representatives, $\bm S$, are first initialized by randomly selecting an instance from each bag. A hyper-plane is then obtained using the representative instances, and new optimal representatives are found with respect to the current hyper-plane, by using the rule given in Equation~\eqref{eq:si}. At each step, the previous values in $\bm S$ are stored in $\bm S_{old}$. The training procedure ends when the bag representatives stop changing from one iteration to the next ($\bm S = \bm S_{old}$). During the testing procedure, each bag produces an output vector based on the hyper-plane found in the training procedure. The bag label is then assigned by taking the sign of the output vector's maximum value, following the SMI assumption.

\begin{figure}
    \centering
    \label{fig:diff}
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{mirsvm_figure.eps} % first figure itself
    \end{minipage}\hfill
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{misvm_figure.eps} % second figure itself
    \end{minipage}
    \caption{Difference between MIRSVM and MISVM on a random $2$-dimensional toy dataset. The instances of this dataset were drawn from the standard normal distribution, with $6$ bags, $4$ of which are negative (red) and $2$ of which are positive (blue), indicated by the red and blue dotted circles. There are a total of $125$ instances, $65$ belonging to the positive bags and $60$ belonging to the negative bags. The decision boundaries (black lines) are shown for each of the algorithms, along with the SVM margin (red and blue dashed lines) and support vectors (black crosses). In this example, both MIRSVM and MISVM were trained using $C = 1000$ and $\sigma = 1.5$. Note the differing number of support vectors produced by the two methods. MIRSVM has $6$, one for each bag, and MISVM has $29$. Also note the smoother representation of the data distribution given by MIRSVM's decision boundary, unlike MISVM whose decision boundary was greatly affected by the larger number of support vectors belonging to the negative class with respect to the only $2$ positive support vectors. }
\end{figure}
\begin{figure}
\centering
\label{fig:convergence}
\includegraphics[width=0.95\textwidth]{convergence.eps} 
\caption{Bag representative convergence plots on 9 datasets. The blue line shows the number of bag representatives that are equal from one iteration to the next. The red dashed line represents the total number of bags.}
\end{figure}

This formulation is designed to utilize and select representatives from positive and negative bags, unlike MISVM, which only optimizes over representatives from positive bags, while flattening the negative bag instances. MISVM allows multiple representatives to be chosen from negative bags and limits positive bag-representatives to be one, while MIRSVM allows for balanced bag-representative selection, where each bag is allowed one. MISVM also uses a wrapper technique to initialize the positive bag-representatives by taking the mean vector of the instances within each positive bag. This is an implicit assumption that the instances within the positive bags are all positive, whereas MIRSVM's initialization procedure selects an instance from all bags at random, ensuring no noise is added by any wrapper techniques during initialization and no assumptions are made about the instances. Due to the constraints on the representatives, MIRSVM produces sparser models while MISVM has the freedom to select as many negative support vectors as it needs and restricts the support vectors chosen from positive bags to be one. Figure~\ref{fig:diff} shows the decision boundaries produced by MIRSVM and MISVM to highlight the differences in their solutions. As Figure~\ref{fig:diff} shows, MISVM produces a larger number of support vectors from the negative bags, which greatly influences the final decision boundary in favor of the negative class.

>>>>

Algorithm~\ref{alg:mirsvm} shows the procedure for training the multi-instance representative SVM classifier and obtaining the optimal representatives from each bag.  During training, the representatives, $\bm S$, are first initialized by randomly selecting an instance from each bag. A hyper-plane is then obtained using the representative instances, and new optimal representatives are found with respect to the current hyper-plane, by using the rule given in Equation~\eqref{eq:si}. At each step, the previous values in $\bm S$ are stored in $\bm S_{old}$. The training procedure ends when the bag representatives stop changing from one iteration to the next ($\bm S = \bm S_{old}$). Examples of the convergence of bag-representatives are shown in Figure~\ref{fig:convegence}. During the testing procedure, each bag produces an output vector based on the hyper-plane found in the training procedure. The bag label is then assigned by taking the sign of the output vector's maximum value, following the SMI assumption.

\begin{figure}
\centering
\label{fig:convegence}
\includegraphics[width=0.95\textwidth]{convergence.eps} 
\caption{Bag representative convergence plots on 9 datasets. The blue line shows the number of bag representatives that are equal from one iteration to the next. The red dashed line represents the total number of bags.}
\end{figure}

\begin{figure}
    \centering
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{mirsvm_figure.eps} % first figure itself
    \end{minipage}\hfill
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{misvm_figure.eps} % second figure itself
    \end{minipage}
    \caption{Difference between MIRSVM and MISVM on a random $2$-dimensional toy dataset. The instances of this dataset were drawn from the standard normal distribution, with $6$ bags, $4$ of which are negative (red) and $2$ of which are positive (blue), indicated by the red and blue dotted circles. There are a total of $125$ instances, $65$ belonging to the positive bags and $60$ belonging to the negative bags. The decision boundaries (black lines) are shown for each of the algorithms, along with the SVM margin (red and blue dashed lines) and support vectors (black crosses). In this example, both MIRSVM and MISVM were trained using $C = 1000$ and $\sigma = 1.5$. Note the differing number of support vectors produced by the two methods. MIRSVM has $6$, one for each bag, and MISVM has $29$. Also note the smoother representation of the data distribution given by MIRSVM's decision boundary, unlike MISVM whose decision boundary was greatly affected by the larger number of support vectors belonging to the negative class with respect to the only $2$ positive support vectors. }\label{fig:diff}
\end{figure}

This formulation is designed to utilize and select representatives from positive and negative bags, unlike MISVM, which only optimizes over representatives from positive bags, while flattening the negative bag instances. MISVM allows multiple representatives to be chosen from negative bags and limits positive bag-representatives to be one, while MIRSVM allows for balanced bag-representative selection, where each bag is allowed one. MISVM also uses a wrapper method to initialize the positive bag-representatives by taking the mean vector of the instances within each positive bag. This is an implicit assumption that the instances within the positive bags are all positive, whereas MIRSVM's initialization procedure selects an instance from all bags at random, ensuring no noise is added by any wrapper techniques during initialization and no assumptions are made about the instances. Due to the constraints on the representatives, MIRSVM produces sparser models while MISVM has the freedom to select as many negative support vectors as it needs and restricts the support vectors chosen from positive bags to be one. Figure~\ref{fig:diff} shows the decision boundaries produced by MIRSVM and MISVM to highlight the differences in their solutions. As Figure~\ref{fig:diff} shows, MISVM produces a larger number of support vectors from the negative bags, which greatly influences the final decision boundary in favor of the negative class.
